% ============================================================
% 20. TEAM CONTRIBUTION
% ============================================================
\chapter{Team Contribution}

This chapter documents the individual contributions of each team member throughout the development of the zGate Zero-Trust Database Access Gateway. The project required expertise across multiple domains including backend systems programming, frontend development, security engineering, and research.

\section{Overview of Contribution Approach}

The zGate project was developed by a seven-member team, with responsibilities distributed across the following primary domains:

\begin{itemize}
    \item \textbf{Backend \& Proxy Development:} Core proxy logic, protocol handlers, connection management
    \item \textbf{Security Engineering:} Authentication, authorization, encryption, and access control
    \item \textbf{Frontend Development:} Web-based administration dashboard (Next.js/React)
    \item \textbf{CLI Development:} Command-line interface for end-user database access
    \item \textbf{Research \& Architecture:} Zero-Trust research, competitive analysis, system design
    \item \textbf{Documentation:} Technical documentation, code documentation, presentation materials
\end{itemize}

Table~\ref{tab:contribution-overview} provides a high-level summary of each member's primary focus areas.

\begin{table}[H]
\centering
\caption{Team Contribution Overview}
\label{tab:contribution-overview}
\begin{tabular}{|l|l|}
\hline
\textbf{Team Member} & \textbf{Primary Focus Areas} \\
\hline
Moustafa Hashem & Team Lead, Core Architecture, Authentication, CLI \\
Hana Shamel & PostgreSQL Proxy, Data Masking, CLI, Research \\
Kareem Ehab & MySQL Proxy, Interceptors, Audit Logging, Security \\
Michael George & Storage Infrastructure, Encryption, Identity Management \\
Mayar Walid & MongoDB Proxy, WebUI Development, API Integration \\
Karen Maurice & Zero-Trust Research, WebUI Development, Theming \\
Rodina Mohamed & MSSQL Protocol, TLS Security, Testing \\
\hline
\end{tabular}
\end{table}

\section{Individual Contributions}

\subsection{Moustafa Hashem -- Team Lead}

As Team Lead, Moustafa directed the full project lifecycle, including sprint planning, task assignment, and cross-team communication. His technical contributions spanned the core system architecture and critical security components.

\subsubsection{Core System Architecture}
\begin{itemize}
    \item Designed the Zero-Trust database proxy architecture, enabling secure identity-based access control and traffic interception.
    \item Built the core networking foundation: dynamic listeners, connection forwarding logic, and high-performance connection pooling.
    \item Developed the unified Handler Interface, ensuring protocol consistency across MySQL, PostgreSQL, and MSSQL implementations.
\end{itemize}

\subsubsection{Dynamic Connection \& Identity Management}
\begin{itemize}
    \item Engineered the dynamic proxy system that creates isolated per-session ports (eliminating static listener vulnerabilities).
    \item Designed multiple identity-management models:
    \begin{itemize}
        \item Temporary database users (ephemeral credentials)
        \item Shared credential pools
        \item User-specific persistent accounts
    \end{itemize}
    \item Implemented both Local Mode tunneling and Remote Mode direct access patterns.
\end{itemize}

\subsubsection{Security, Authentication \& RBAC}
\begin{itemize}
    \item Built the complete JWT authentication system, supporting login, token refresh, revocation, and secure session tracking.
    \item Designed the RBAC policy engine, enabling granular permissions at the database and table levels.
\end{itemize}

\subsubsection{CLI \& Admin Tools}
\begin{itemize}
    \item Re-engineered the CLI architecture to improve security, consistency, and maintainability.
    \item Developed admin tools for user provisioning, credential management, and session visibility.
\end{itemize}

\subsubsection{System Reliability}
\begin{itemize}
    \item Engineered the core token lifecycle logic, ensuring consistent expiration handling and safe session cleanup.
    \item Designed system reliability mechanisms for stable startup, shutdown, and error recovery across components.
\end{itemize}

%-----------------------------------------------------------

\subsection{Hana Shamel}

Hana contributed to proxy development, CLI implementation, and conducted extensive research on Zero-Trust security models.

\subsubsection{Proxy \& Backend Development}
\begin{itemize}
    \item Developed the initial PostgreSQL proxy prototype, establishing early connectivity and packet-flow behavior.
    \item Implemented the data masking engine used to redact sensitive information from database responses.
    \item Contributed to backend architecture and low-level packet processing design.
\end{itemize}

\subsubsection{CLI Development}
\begin{itemize}
    \item Built essential CLI authentication and database command workflows.
    \item Integrated OS-native secure credential storage using:
    \begin{itemize}
        \item macOS Keychain
        \item Windows Credential Manager
        \item Linux Secret Service
    \end{itemize}
\end{itemize}

\subsubsection{Security \& Research}
\begin{itemize}
    \item Conducted research on Zero-Trust models, database attack vectors, and secure-access frameworks.
    \item Investigated observability and monitoring tools to support long-term system stability.
    \item Studied industry attack statistics and market needs to guide feature development.
\end{itemize}

\subsubsection{Documentation}
\begin{itemize}
    \item Wrote documentation for SQL masking logic, CLI behavior, and proxy components.
    \item Contributed to PowerPoint presentation materials and project documentation structure.
\end{itemize}

%-----------------------------------------------------------

\subsection{Kareem Ehab}

Kareem led the MySQL proxy implementation and developed the security interceptor framework central to the system's query protection capabilities.

\subsubsection{MySQL Proxy \& Backend Development}
\begin{itemize}
    \item Implemented the full MySQL proxy logic, including authentication handshake, packet forwarding, and controlled shutdown.
    \item Built the extensible interceptor framework used for data masking, SQL safety checks, and logging.
\end{itemize}

\subsubsection{Security Enforcement}
\begin{itemize}
    \item Implemented the Query Safety Enforcement Engine, blocking malicious or destructive SQL statements in real time.
    \item Integrated TLS/SSL certificate verification to secure proxy connections.
    \item Added secure temporary credential logic to handle session-specific identities.
\end{itemize}

\subsubsection{Audit Logging \& Observability}
\begin{itemize}
    \item Developed the Comprehensive Audit Logging System, capturing:
    \begin{itemize}
        \item Query history with user attribution
        \item Authentication events
        \item Proxy metrics
        \item Session activity
    \end{itemize}
    \item Added internal monitoring endpoints to enhance observability and debugging capabilities.
\end{itemize}

\subsubsection{Research \& Documentation}
\begin{itemize}
    \item Conducted research on database threats, competitor proxy architectures, and security standards.
    \item Contributed technical flow explanations and feature breakdowns for presentation materials.
\end{itemize}

%-----------------------------------------------------------

\subsection{Michael George}

Michael focused on storage infrastructure, encryption systems, and identity management backend components.

\subsubsection{Proxy \& Early Prototype Development}
\begin{itemize}
    \item Developed the early Python-based proxy prototypes to validate networking concepts.
    \item Implemented the initial PostgreSQL proxy prototype before transitioning to the Go-based design.
\end{itemize}

\subsubsection{Storage \& Encryption Infrastructure}
\begin{itemize}
    \item Migrated system configuration from YAML to SQLite, enabling persistent structured storage for users, roles, tokens, and database configurations.
    \item Implemented AES-256-GCM encryption for credentials, admin passwords, and sensitive configurations.
\end{itemize}

\subsubsection{Identity \& Access Management}
\begin{itemize}
    \item Implemented backend logic for:
    \begin{itemize}
        \item User account creation and update
        \item Password hashing and verification
        \item Role assignment and permission control
    \end{itemize}
    \item Added full CRUD operations for role management.
\end{itemize}

\subsubsection{Documentation}
\begin{itemize}
    \item Wrote documentation for storage architecture, encryption system, and user/role backend features.
\end{itemize}

%-----------------------------------------------------------

\subsection{Mayar Walid}

Mayar contributed to MongoDB proxy development and led significant portions of the WebUI development, particularly monitoring and dashboard interfaces.

\subsubsection{Proxy \& Backend Development}
\begin{itemize}
    \item Built the initial MongoDB proxy prototype supporting early database routing tests.
    \item Implemented backend endpoints for Admin Account Management (CRUD operations).
\end{itemize}

\subsubsection{Web UI Development (Next.js \& React)}
\begin{itemize}
    \item \textbf{Shared \& Personal Database Accounts:} Developed UI interfaces for shared credential pools and user-specific account management with validation, search filters, and interactive modals.
    \item \textbf{Monitoring \& Query Interfaces:}
    \begin{itemize}
        \item Session Monitoring Dashboard with real-time session visibility
        \item Query History Viewer with dynamic updates, filters, and search tools
    \end{itemize}
    \item \textbf{Dashboards:} Developed the Overview Dashboard displaying live system metrics and enhanced session, query, and database pages with improved functionality.
\end{itemize}

\subsubsection{API Integration \& State Management}
\begin{itemize}
    \item Integrated UI components with backend APIs for authentication, roles, permissions, databases, sessions, and admin operations.
    \item Ensured synchronized application state across the UI.
\end{itemize}

\subsubsection{Documentation}
\begin{itemize}
    \item Contributed UI visuals, workflow explanations, and dashboard breakdowns for presentation slides.
\end{itemize}

%-----------------------------------------------------------

\subsection{Karen Maurice}

Karen conducted Zero-Trust research and contributed extensively to WebUI development, including theming systems and access control interfaces.

\subsubsection{Zero-Trust Research \& Prototype Engineering}
\begin{itemize}
    \item Conducted research into Zero-Trust architecture, database security, and access-control models.
    \item Built the Zero-Trust gateway prototype using Go and Docker.
    \item Designed the zGate Secure Pipeline Architecture, featuring:
    \begin{itemize}
        \item mTLS (mutual TLS) security layer
        \item JSONL protocol layer
        \item Security processing pipeline
        \item Strategy adapter layer for database engines
    \end{itemize}
\end{itemize}

\subsubsection{Web UI Development (Next.js \& React)}
\begin{itemize}
    \item \textbf{User, Role \& Database Management:} Developed UI systems for user, role, database, and access-control management with permission indicators and validation layers.
    \item \textbf{UI/UX Systems:}
    \begin{itemize}
        \item Implemented light/dark mode theming
        \item Built advanced UI elements including dialogs, permission selectors, detailed views, and filters
    \end{itemize}
    \item \textbf{Monitoring Interfaces:} Developed Session Monitoring Dashboard and Query History Viewer with filtering and search capabilities.
\end{itemize}

\subsubsection{API Integration \& State Management}
\begin{itemize}
    \item Integrated Web UI with backend APIs for authentication, roles, databases, and sessions.
    \item Implemented robust state management and error-handling flows.
\end{itemize}

\subsubsection{Documentation}
\begin{itemize}
    \item Contributed to architectural documentation, UI flow diagrams, and frontend integration notes.
\end{itemize}

%-----------------------------------------------------------

\subsection{Rodina Mohamed}

Rodina focused on MSSQL protocol development, security infrastructure, and testing validation.

\subsubsection{MSSQL Protocol Development}
\begin{itemize}
    \item Implemented the initial MSSQL proxy framework, including support for the TDS (Tabular Data Stream) protocol.
    \item Developed foundational packet-handling logic for Microsoft SQL Server communication.
\end{itemize}

\subsubsection{Security \& Encryption Infrastructure}
\begin{itemize}
    \item Contributed to the design of the TLS/SSL encryption layer, securing proxy communication.
    \item Participated in building engine-agnostic secure protocol abstractions.
\end{itemize}

\subsubsection{Research \& Architectural Input}
\begin{itemize}
    \item Performed research into:
    \begin{itemize}
        \item Zero-Trust architecture principles
        \item Market security needs and requirements
        \item Relevant software design patterns
    \end{itemize}
    \item Provided architectural feedback improving system modularity and scalability.
\end{itemize}

\subsubsection{Testing \& Validation}
\begin{itemize}
    \item Built test applications to verify MSSQL proxy behavior.
    \item Performed multi-tool validation for stability and correctness.
\end{itemize}

\subsubsection{Documentation}
\begin{itemize}
    \item Documented MSSQL protocol behavior, TDS packet flow, security considerations, and integration architecture for MSSQL components.
\end{itemize}
